Building Markov State Model for a Periodically Driven Non-Equilibrium System

### 2014-08-28 writedown/period.noneq.msm/main.tex
backup before merge with C.S.

The alanine dipeptide was put into the local thermostating
environment, and was driven by a periodic electric field with period
$T = 10$~ps. The 20,000 branching trajectories started from 20,000
initial configurations that samples the equilibrium distribution.
\redc{Write a lot of details on non-equilibrium MD.}  The branching
trajectories were each 1000~ps long, and the system reaches
non-equilibrium steady state in roughly 300~ps. We compute the
generator from non-equilibrium trajectories in time interval $[t_1,
t_2]$. If the reduced dynamics were periodic, then generator $\vect
M(t)$ would not depends on the choice of the time interval. We compute
the reduced generator by two discretizations $K=2$ and $K=20$, and two
choices of time intervals $[0, 40]$ and $[320, 1000]$, and then
compare the results with the brute force result in
Fig.~\ref{fig:tmp2}. In the figure, we show the time-dependent
probability $\mathbb P\big(\phi\in[0,180), \psi\in [0,180)\big)$.
Using $K=2$ the dynamics depends on the time interval used for
calculating the generator: using time intervals $[0, 40]$ the reduced
dynamics can only reproduces the brute force result in $[0, 40]$,
while using time intervals $[0, 40]$ the reduced dynamics can only
reproduces the brute force result after 300~ps.  This shows that the
reduced dynamics is lack of periodicity.  The reason is that the
discretization is too coarse so that the dynamics cannot be fully
equilibriated within the lag-time $\tau$ in each discretized set.  For
$K=20$, the reduced dynamics does not depend on the time interval of
calculating the generator, and is consistent with the brute force
non-equilibrium simulation within the error bar. Therefore, the $K=20$
is a reliable discretization for the alanine dipeptide system.

@@

\subsection{Core set identification}

@@

The coresets are denoted by $C_{R_\alpha}$, $C_\beta$ and $C_{L_\alpha}$, which correspond to
right-handed alpha-helix, beta-sheet and left-handed alpha-helix, respectively.

\begin{figure}
  \centering  
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-dist.eps}
  % \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-dist-msm.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-floquet-vec-1.eps}
  \caption{The steady state distribution calculated by brute force non-equilibrium comparing with MSM.}
  \label{fig:num-1}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-msm-1.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-msm-1.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-msm-1.eps}
  \caption{The forward and backward committors computed by MSM is compared with those computed by brute force non-equilibrium simulations.}
  \label{fig:num-3}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-msm-2.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-msm-2.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-msm-2.eps}
  \caption{The forward and backward committors computed by MSM is compared with those computed by brute force non-equilibrium simulations.}  
  \label{fig:num-4}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-msm-3.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-msm-3.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-msm-3.eps}
  \caption{The forward and backward committors computed by MSM is compared with those computed by brute force non-equilibrium simulations.}  
  \label{fig:num-5}
\end{figure}

\section{Discussions}
\subsection{The periodicity of the reduced dynamics}

As discussed before, if the reduced dynamics captures the slow
dynamics of interest, then the reduced dynamics is also periodic,
which implies that the generator $M(t)$ can be estimated from any time
interval by Eq.~\eqref{eqn:tmp4}. In practice, the reduced dynamics
may not be perfectly periodic. The generator calculated from
different time intervals can be used to test the periodicity of the
reduced dynamics. The results are shown in Tab.~\ref{tab:tmp1} and
Fig.~\ref{fig:tmp2}.  The dependency with respect to the time
interval, in which the generator is calculated, indicates the reduced
dynamics is not perfectly periodic. The reduced dynamics from $[0,40]$
reproduces the best probability in interval $[0,40]$, and worst in the
steady state. Moreover, the leading non-trivial timescale is also off
(see Tab.~\ref{tab:tmp1}), while the accuracy of the second and third
non-trivial timescales are better, because they are comparable to the
period, 40~ps, in which the dynamics is relaxed. The reduced dynamics
from $[300,1000]$ reproduces the best steady state probability, while
worst in reproducing the initial relaxation process. The reduced
dynamics from $[0,320]$ is indistinguishable from the one from
$[300,1000]$, which implies that in practice, one only have to
simulation relatively short non-equilibrium trajectories (that are
almost relaxed), and compute the generator from it to build the reduced dynamics. \redc{This saves a great
amount of computational resources comparing with calculating the
reduced dynamics only from steady state trajectories.}

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{fig/t040/fig-cg-prob-range.eps}  
  \caption{The time-dependent probability of conformation
    $\confc$. The brute force MD simulation is compared with reduced
    dynamics, the transition matrix of which are calculated from
    different ranges. The red shadow region is the statistical uncertainty of the brute force MD simulation.
    The inserts show the zoom-ined probability at the relaxation stage $[0,120 ]~\textrm{ps}$ and at the steady state $[800,880 ]~\textrm{ps}$.}
  \label{fig:tmp2}
\end{figure}

\subsection{The choice of lag time $\tau$}
Table~\ref{tab:tmp1} shows a clear trend of worse accuracy with
increased $\tau$, however, the timescales are not very sensitive to
the choice of lag time $\tau$.

\section{To do}

\begin{itemize}
\item Theoretical analysis of the accuracy of the discretization, a
generalization of the equilibrium case. the numerical results indicate
a very close link. The perodicity is very nice, if the generator is
``continuous'', then it should be uniformly continuous, then it should
be possible to develop estimates that are similar to the equilibrium
case.
\item More tests on the effect of choosing different lag-time $\tau$ (0.5 and 1.0~ps are done).
\item More theoretical consideration on the Floquet theory, what is the
explict form of matrix $\vect B$, and does it has any nice properties,
such as self-joint.
\end{itemize}

\section{Reversibility of the original dynamics}

@@

>>>>

The alanine dipeptide was put into the local thermostating
environment, and was driven by a periodic electric field with period
$T = 10$~ps. The 20,000 branching trajectories started from 20,000
initial configurations that samples the equilibrium distribution.
\redc{Write a lot of details on non-equilibrium MD.}  The branching
trajectories were each 1000~ps long, and the system reaches
non-equilibrium steady state in roughly 300~ps. We compute the
generator from non-equilibrium trajectories in time interval $[t_1,
t_2]$. If the reduced dynamics were periodic, then generator $\vect
M(t)$ would not depends on the choice of the time interval. We compute
the reduced generator by two discretizations $K=2$ and $K=20$, and two
choices of time intervals $[0, 40]$ and $[320, 1000]$, and then
compare the results with the brute force result in
Fig.~\ref{fig:tmp2}. In the figure, we show the time-dependent
probability $\mathbb P\big(\phi\in[0,180), \psi\in [0,180)\big)$.
Using $K=2$ the dynamics depends on the time interval used for
calculating the generator: using time intervals $[320, 1000]$ the reduced
dynamics can only reproduces the brute force result in $[0, 40]$,
while using time intervals $[0, 40]$ the reduced dynamics can only
reproduces the brute force result after 300~ps.  This shows that the
reduced dynamics is lack of periodicity.  The reason is that the
discretization is too coarse so that the dynamics cannot be fully
equilibriated within the lag-time $\tau$ in each discretized set.  For
$K=20$, the reduced dynamics does not depend on the time interval of
calculating the generator, and is consistent with the brute force
non-equilibrium simulation within the error bar. Therefore, the $K=20$
is a reliable discretization for the alanine dipeptide system.

@@

\begin{figure}
  \centering  
  \includegraphics[width=0.4\textwidth]{fig/t010/cluster.marco.40/fig-dist.eps}
  % \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-dist-msm.eps}
  \includegraphics[width=0.4\textwidth]{fig/t010/cluster.marco.40/fig-floquet-vec-1.eps}
  \caption{The steady state distribution calculated by brute force non-equilibrium comparing with MSM.}
  \label{fig:num-1}
\end{figure}

\subsection{Coreset identification}

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{fig/t040/cluster.marco/fig-cluster-2.eps}
  \caption{The coreset identification. Different colors indicate different coresets: $C_{R_\alpha}$ (yellow), $C_\beta$ (red) and $C_{L_\alpha}$ (green).
    The blue color means out of any coreset. (By Marco)}
  \label{fig:cluster}
\end{figure}

@@

The coresets (see Fig.~\ref{fig:cluster}) are denoted by $C_{R_\alpha}$ (yellow), $C_\beta$ (red) and $C_{L_\alpha}$ (green), which correspond to
right-handed alpha-helix, beta-sheet and left-handed alpha-helix, respectively.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-1.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-msm-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-msm-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-msm-1.eps}
  \caption{The forward $q^\fwd_{L_\alpha}$ and backward committors $q^\bwd_{L_\alpha}$ computed by MSM (second row) is compared with those computed by brute force non-equilibrium simulations (first row).}
  \label{fig:num-3}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-2.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-msm-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-msm-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-msm-2.eps}
  \caption{The forward $q^\fwd_{R_\alpha}$ and backward $q^\fwd_{R_\alpha}$ committors computed by MSM (second row) is compared with those computed by brute force non-equilibrium simulations (first row).}  
  \label{fig:num-4}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-3.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-fw-msm-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-bw-msm-3.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40/fig-commitor-diff-msm-3.eps}
  \caption{The forward $q^\fwd_{\beta}$ and backward $q^\fwd_{\beta}$ committors computed by MSM (second row) is compared with those computed by brute force non-equilibrium simulations (first row).}  
  \label{fig:num-5}
\end{figure}

\subsection{First mean hitting time}

The first mean hitting time is presented in Fig.~\ref{fig:num-6}.

\begin{figure}
  \centering
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40.fht/fig-fht-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40.fht/fig-fht-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40.fht/fig-fht-3.eps}\\
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40.fht/fig-fht-msm-1.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40.fht/fig-fht-msm-2.eps}
  \includegraphics[width=0.23\textwidth]{fig/t010/cluster.marco.40.fht/fig-fht-msm-3.eps}
  \caption{The first mean htting time comparsion between the brute force simulation (first row) and the MSM (second row).
  From left to right are first mean hitting time to coresets $C_{L_\alpha}$, $C_{R_\alpha}$ and $C_{\beta}$, respectively}  
  \label{fig:num-6}
\end{figure}

% \section{Discussions}
% \subsection{The periodicity of the reduced dynamics}

% As discussed before, if the reduced dynamics captures the slow
% dynamics of interest, then the reduced dynamics is also periodic,
% which implies that the generator $M(t)$ can be estimated from any time
% interval by Eq.~\eqref{eqn:tmp4}. In practice, the reduced dynamics
% may not be perfectly periodic. The generator calculated from
% different time intervals can be used to test the periodicity of the
% reduced dynamics. The results are shown in Tab.~\ref{tab:tmp1} and
% Fig.~\ref{fig:tmp2}.  The dependency with respect to the time
% interval, in which the generator is calculated, indicates the reduced
% dynamics is not perfectly periodic. The reduced dynamics from $[0,40]$
% reproduces the best probability in interval $[0,40]$, and worst in the
% steady state. Moreover, the leading non-trivial timescale is also off
% (see Tab.~\ref{tab:tmp1}), while the accuracy of the second and third
% non-trivial timescales are better, because they are comparable to the
% period, 40~ps, in which the dynamics is relaxed. The reduced dynamics
% from $[300,1000]$ reproduces the best steady state probability, while
% worst in reproducing the initial relaxation process. The reduced
% dynamics from $[0,320]$ is indistinguishable from the one from
% $[300,1000]$, which implies that in practice, one only have to
% simulation relatively short non-equilibrium trajectories (that are
% almost relaxed), and compute the generator from it to build the reduced dynamics. \redc{This saves a great
% amount of computational resources comparing with calculating the
% reduced dynamics only from steady state trajectories.}

% \begin{figure}
%   \centering
%   \includegraphics[width=0.5\textwidth]{fig/t040/fig-cg-prob-range.eps}  
%   \caption{The time-dependent probability of conformation
%     $\confc$. The brute force MD simulation is compared with reduced
%     dynamics, the transition matrix of which are calculated from
%     different ranges. The red shadow region is the statistical uncertainty of the brute force MD simulation.
%     The inserts show the zoom-ined probability at the relaxation stage $[0,120 ]~\textrm{ps}$ and at the steady state $[800,880 ]~\textrm{ps}$.}
%   \label{fig:tmp2}
% \end{figure}

% \subsection{The choice of lag time $\tau$}
% Table~\ref{tab:tmp1} shows a clear trend of worse accuracy with
% increased $\tau$, however, the timescales are not very sensitive to
% the choice of lag time $\tau$.

% \section{To do}

% \begin{itemize}
% \item Theoretical analysis of the accuracy of the discretization, a
% generalization of the equilibrium case. the numerical results indicate
% a very close link. The perodicity is very nice, if the generator is
% ``continuous'', then it should be uniformly continuous, then it should
% be possible to develop estimates that are similar to the equilibrium
% case.
% \item More tests on the effect of choosing different lag-time $\tau$ (0.5 and 1.0~ps are done).
% \item More theoretical consideration on the Floquet theory, what is the
% explict form of matrix $\vect B$, and does it has any nice properties,
% such as self-joint.
% \end{itemize}

\section{Approximation quality of the Markov state model}

The whole Floquet theory can be applied to the original dynamics, and we denote that the resulting transition matrix by $\psi(T)$, then
the time-dependent probability density can be written as
\begin{align*}
  \rho(\vect x, kT) = 
\end{align*}

In general, the original dynamics of the molecular system under perodical external driven
is not reversible, see the discussion in Sec.~\ref{sec:revs}. However, we can assume the,

\section{Reversibility of the original dynamics}
\label{sec:revs}

@@

\subsection{Irreversibility of the perodical symmetrical dynamics}
